A weakly structured stem for human origins in Africa

### 2022-07-22 paper.tex

Regardless of the model choice for early epochs, inference of human demographic
history for the last 150ka was remarkably robust.
In a reticulated model, we use ``divergence'' between populations to describe
the time of their most recent shared ancestry.
The earliest divergence among
contemporary human populations differentiates the southern African Nama from
other African groups between 110--135ka, with low to moderate levels of
subsequent gene flow (Table \ref{tab:migration-rates}). In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$140ka. 
We conclude that geographic patterns of contemporary \emph{Homo sapiens} population structure
date back to the late Middle Stone Age in Africa, likely arising during MIS 5.
Although we find evidence for earlier population structure in Africa (see
below), contemporary populations cannot be easily mapped onto the more ancient
`stem' groups as only a small proportion of drift between contemporary populations can
be attributed to drift between stems (Figures~\ref{fig:predictions} and
\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).

>>>>

Regardless of the model choice for early epochs, inference of human demographic
history for the last 150ka was remarkably robust.
In a reticulated model, we use ``divergence'' between populations to describe
the time of their most recent shared ancestry.
The earliest divergence among
contemporary human populations differentiates the southern African Nama from
other African groups between 110--135ka, with low to moderate levels of
subsequent gene flow (Table \ref{tab:migration-rates}). In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$140ka. 
We conclude that geographic patterns of contemporary \emph{Homo sapiens} population structure
date back to the late Middle Stone Age in Africa, likely arising during MIS 5.
Although we find evidence for earlier population structure in Africa (see
below), contemporary populations cannot be easily mapped onto the more ancient
`stem' groups as only a small proportion of drift between contemporary populations can
be attributed to drift between stems (Section \ref{sec:f4},  Figures~\ref{fig:predictions} and
\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).
